---
title: "Selection of 10 Klebsiella pneumoniae genomes for analysis"
author: "Eden Black"
output: html_document
---
```{r message=FALSE}
library(tidyverse)
```


candidates.tsv was assembled from the NCBI database searching for complete Klebsiella
pneumoniae genomes from RefSeq using the following code in the NCBI CLI:

datasets summary genome taxon "Klebsiella pneumoniae" \
  --assembly-level complete --assembly-source refseq \
  --as-json-lines \
| dataformat tsv genome --fields \
accession,source_database,organism-name,organism-infraspecific-strain,\
assminfo-level,assmstats-total-sequence-len,assmstats-number-of-contigs,\
assmstats-total-number-of-chromosomes,assmstats-gc-percent,\
checkm-completeness,checkm-contamination,assminfo-sequencing-tech,\
assminfo-biosample-accession,assminfo-biosample-isolation-source,\
assminfo-biosample-host,assminfo-biosample-host-disease,\
assminfo-biosample-geo-loc-name,assminfo-biosample-collection-date,\
assminfo-bioproject,assminfo-bioproject-lineage-title,assminfo-release-date \
> data/candidates.tsv

As a case study for this minimal analysis, the aim is to analyse the genomes of different 
samples from a single geographical location to identify similarities and differences,
particularly intra- and inter-strain.

```{r data-import}
candidates = read_tsv('data/candidates.tsv') %>%
    mutate(`Assembly BioSample Accession` = as.factor(`Assembly BioSample Accession`))
colnames(candidates)
```

```{r locations}
candidate_locations = candidates %>%
    group_by(`Assembly BioSample Geographic location`) %>%
    summarise(count = n()) %>%
    filter(count >= 10) %>%
    arrange(desc(count))

head(candidate_locations, 10)
```

While many samples have their locations logged only by country, two cities appear on the
top ten list of locations: Seoul (85) and Nanchang (83).

``` {r qc}
seoul_candidates = candidates %>%
    filter(`Assembly BioSample Geographic location` == 'South Korea: Seoul' &
    `Assembly Stats Number of Contigs` <= 10 &
    `CheckM completeness` >= 98 &
    `CheckM contamination` < 1)

nrow(seoul_candidates)
seoul_candidates %>% count(`Assembly BioProject Accession`, sort = TRUE)
sprintf('For samples from Seoul, the mean completeness is %.1f and mean contamination is %.1f',
    mean(seoul_candidates$`CheckM completeness`),
    mean(seoul_candidates$`CheckM contamination`))

nanchang_candidates = candidates %>%
    filter(`Assembly BioSample Geographic location` == 'China:Nanchang' &
    `Assembly Stats Number of Contigs` <= 10 &
    `CheckM completeness` >= 98 &
    `CheckM contamination` < 1)

nrow(nanchang_candidates)
nanchang_candidates %>% count(`Assembly BioProject Accession`, sort = TRUE)
sprintf('For samples from Nanchang, the mean completeness is %.1f and mean contamination is %.1f',
    mean(nanchang_candidates$`CheckM completeness`),
    mean(nanchang_candidates$`CheckM contamination`))
```

71 candidates from Nanchang passed QC, compared to only 10 from Seoul. As such, the samples
from Nanchang were taken forward.

The final step for candidate selection was sample independence. Genomes submitted in the
same project are more likely to be the same or very closely-related isolate 
sampled repeatedly. As 28 samples were submitted as individuals, they were selected to be 
taken forward for multilocus sequence typing.

```{r nanchang-filter}
nanchang_candidates = nanchang_candidates %>%
    filter(!(`Assembly BioProject Accession` %in% c('PRJNA1474683', 'PRJNA672246',
    'PRJNA1469284', 'PRJNA1344441')))

nrow(nanchang_candidates)
```

Finally, the accession numbers of the candidate genomes were exported to a .txt file to
download from NCBI.

```{r write-txt}
write_lines(nanchang_candidates$`Assembly Accession`, 'data/accession.txt')
```