---
title: "Analysis of prophages in 10 Klebsiella pneumoniae genomes"
author: "Eden Black"
output: 
  github_document:
    html_preview: true
    df_print: kable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 4,
                      fig.align = "center")

library(tidyverse)
library(knitr)
library(here)

theme_set(theme_minimal(base_size = 11))
theme_update(
  panel.grid.minor = element_blank(),
  plot.title       = element_text(face = "bold")
)

here::i_am("notebooks/final_analysis.Rmd")
```

10 *Klebsiella pneumoniae* genomes isolated in Nanchang, China were chosen for analysis. 6 of the genomes are from the dominant ST11 strain (which accounts for 60% of infections
in China (SOURCE)) and one each of the other four strain types with NCBI entries for isolates from Nanchang. With one genome per lineage for these STs, statistical testing is
not plausible - this is instead a descriptive analysis.

```{r read-tables}
samples <- read_tsv(here("results", "tables", "samples.tsv"))
prophages <- read_tsv(here("results", "tables", "prophages.tsv"))

# First 1000 rows of CARD are N/A, so read_tsv assigns boolean type by default
# Fix with guess_max = Inf
prophage_genes <- read_tsv(here("results", "tables", "prophage_genes.tsv"),
                           guess_max = Inf)
defence_systems <- read_tsv(here("results", "tables", "defense_systems.tsv"))
defence_genes <- read_tsv(here("results", "tables", "defense_genes.tsv"))

# Treat strain as a factor and order by lineage
samples <- samples |>
  mutate(st = factor(paste0("ST", st))) |>
  arrange(st, sample) |>
  mutate(sample = fct_inorder(sample))

#Table for looking up strain of each sample
st_lookup <- select(samples, sample, st)
```

```{r prophage-burden}
prophage_size <- prophages |>
  group_by(sample) |>
  summarise(prophage_bp = sum(length), .groups = "drop")

burden <- samples |>
  left_join(prophage_size, by = "sample") |>
  select(sample, st, n_prophages, n_prophage_cds, prophage_bp)

burden
```
Table 1: Summary of the sequence type and prophage levels of each *K. pneumoniae* genome

```{r burden-chart}
burden_long <- burden |>
  select(sample, st, `Number of prophages` = n_prophages,
         `Prophage sequence length (kb)` = prophage_bp) |>
         pivot_longer(c(`Number of prophages`, `Prophage sequence length (kb)`))


p_burden <- ggplot(burden_long, aes(x = fct(sample), y = value, fill = st)) +
  geom_col() +
  facet_wrap(~name, scales = "free") +
  labs(title = "ST11 genomes have a higher prophage burden",
    subtitle = "Ten complate K. pneumoniae genomes; prophages called with geNomad",
    x = NULL, y = NULL, fill = "Sequence type"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.4, vjust = 0.6))

p_burden
```
Figure 1: Bar chart of the number of prophages (left) and the combined sequence length of prophages (right) in each *K. pneumoniae* genome

The plots show an expected strong correlation between the number of prophages in a bacterial genome and the sequence length of the prophages. More interestingly, all isolates of
the ST11 sequence type contained as many or more integrated prophages as the ST86 isolate, which contained the most prophages of any of the other sequence types. ST11 is both
highly virulent and frequently displays AMR characteristics (SOURCE), both traits often encoded by prophage morons. To investigate, the prophage gene information from pharokka
was examined.

```{r gene-categorisation}
categories <- prophage_genes |>
  mutate(category = replace_na(category, "unknown function"),
         category = as.factor(category)) |>
  count(category, sort = TRUE) |>
  mutate(percent = round(100 * n / sum(n), 1))

categories
```
Table 2: Prophage gene categories

```{r categories-chart}
p_categories <- ggplot(categories, aes(y = fct_rev(category), x = n,
                                       fill = category == "unknown function")) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue")) +
  labs(title = "43.9% of prophage genes have no known function",
       subtitle = sprintf("%d CDS across %d prophages; annotated with Pharokka (PHROG categories)",
                          nrow(prophage_genes), nrow(prophages)),
       y = NULL, x = "Number of CDS") +
  theme(axis.text.y = element_text(angle = 45, hjust = 1, vjust = 1,
                                   size = 6))

p_categories
```
Figure 2: Bar chart of the number of genes across the 10 combined genome assigned to each category by PHROG

